### 2020-06-18 submission/response_to_reviewers.md

**Reviewer #1**

**The authors put together an interesting 10 simple rules article sharing their experiences with code clubs.**

**I find the proposed code clubs an interesting idea and would be interested in trying this out in the lab. However, quite a bit more practical advice from this manuscript would be helpful to get up and running. Practical advice can be in the form of things that worked, but also explicitly mentioning what didn't work so that others can learn from the authors experience.**

> Thank you for the feedback on where the manuscript could provide more context. Throughout their review, the reviewer encourages us to make more comments about groups where there are more people with CS backgrounds. Admittedly, our lab is made of researchers with non-CS backgrounds and so our implementation of Code Clubs is with that audience in mind. We have added text as appropriate to respond to the following suggestions.

**Some suggestions:**
**Put this into context of the literature: is there any formal description of code clubs? What other learning approaches are commonly used to introduce and improve coding skills in people with no formal programming training?**

**How does this connect to other activities such as hackathons, pair-programming (buddy programming)?**

> We have provided a broader context in the second paragraph of the Introduction and with an expanded discussion of pair-programming in Rule 8.

**Highlight more the benefits of both junior and senior coders, for now the benefits appears to be solely for the junior coders. What motivates the senior folks to join? Or alternatively, in the context of a more diverse lab, how would this work with people with CS and non-CS backgrounds?**

> We have added text to Rule 3 describing the development of more challenging problems for more experienced group members. There is also now a longer section explicitly stating the benefits to more experienced group members and ways to keep their attention by developing sessions around packages and concepts that are new and accessible to everyone.

**Some suggestions on how to implement this would be nice, i.e. how to get started, how to get people motivated to join, etc.**

> It is challenging to comment on this point since we have only done this a part of our normal lab meetings where lab members are compelled to participate. We have added some text to the Conclusion paragraph describing why this model should be successful for more independent groups.

**Are these code clubs unique to biologists? How would this set up work in a hybrid cs/bio environment or perhaps even in a non-bio environment?**

> There's no reason to think that Code Clubs would be unique to biologists. We have added text to the end of the Introduction indicating that as long as the Rules are followed it should be generalizable to groups that are more proficient at programming. Just as Journal Clubs are valuable to all fields of science and levels of experience, we think Code Clubs should also be valuable to the same diversity of participants.

**I would be interested to read more about how much effort it takes in reality from both participants and presenters to prepare? How many hours each week need to be dedicated?**

> In the second to last paragraph of the Introduction we added a sentence indicating that we have found that it takes about as much effort to put together material for a Code Club as it does for a Journal Club session.

**It would be interesting to see a much more extensive proposed program. The examples in Table 1 are illustrative, but far from complete. Can the list be expanded with repeatable activities (git was mentioned in the text) that could constitute a rotating 1 year program and a list of example side activities that are more one-off.**

> Our goal in this manuscript is not to be proscriptive about what the session should contain. We are presenting a framework that diverse groups can use to achieve their own goals with their interests and level of background. We suspect one could certainly create a "schedule" of topics, but the challenge would be sustaining interest across years with the same cohort of participants.

**It would be nice if the above suggestions can be embedded within the rules. To create the space, I think rule 7 and 8 can be merged. Rule 7 seems rather obvious. Rule 6 and 9 are connected and would make sense to have them subsequently, or alternatively as a single rule with two sides of the coin.**

> In our experience, Rule 7 was not obvious and led to the deeper discussion that caused us to re-form our Code Club format. Rather than moving Rules around, we've incorporated the reviewer's comments throughout the manuscript.

**Reviewer #2**
**The manuscript from Hagan, Schloss et al. presents a guide for developing computational skills in labs/groups/buildings/departments, and especially in those environments that include wet lab scientists. This is a timely piece given both the increased importance of computational analyses over the past 1-2 decades, as well as the more recent shuttering of many wet labs due to COVID-19. The piece describes steps to take to develop 1-hour "Code Club" sessions. At these sessions, presenters pose a problem and pair students to solve the problem. After the pair programming, there is a opportunity to reconvene and present solutions.**

**Two general scenarios are explored -- one in which the presenters are helping to solve a problem for the presenter, and another in which the problem is didactic and the key beneficiaries are the other participants. Specific examples for sessions are provided, and additional guidance is offered throughout the rules. Overall, the rules are well organized and the direction is clear.**

**It would appear that the authors violated the first rule of "Code Club" in writing about it, and that is to the benefit of all of us. I have only minor comments:**

> We thank the reviewer for their comments on the manuscript and suggestions for where to add more detail and where we can clarify the text.

**L81-91 (Rule 3): Can you provide more detail in this section? What are examples here of how to "experiment", or what should be avoided? Additionally, returning to basics for new participants seems to conflict with the message to try something new.**

> We have added examples of how we've been experimenting with format during the COVID-19 pandemic to the Introduction as well as variation in difficulty of problems that are assigned in Rule 3.

**L125 (Rule 6): Are the authors recommending a separate GitHub repository for each club meeting, or one general one for the club? Are any of these materials public that they could be shared here as an example? I reviewed the materials at "https://www.riffomonas.org/code_club" but I assume these are more elaborate than what is used in-house.**

> We have added text to Rule 6 detailing what we have found to be pros and cons of hosting material on GitHub. Although interacting with git and GitHub can seem trivial to more experienced people, it can be intimidating to beginners. In many cases, using a dedicated Slack channel or email has lower barriers and allows us to focus on the topic for the session rather than worrying about git configurations or GitHub permissions. Although not all of the examples in Table 1 have repositories on our group's GitHub account, we have included links to those that are available. Looking at these, one will see the variation in "polish" across the sessions, but will also get a sense of the overall approach we have taken to Code Club sessions.

**L32: Suggest rewording "and resulted" to ", resulting".**

> We have edited the text as suggested.

**L60: Suggest rewording "last two for participants." to "last two for non-presenters."**

> We have edited the text as suggested.

### 2020-06-18 submission/manuscript.Rmd
Pat's first take at revision

For most biologists, the ability to generate data has outpaced the ability to analyze those data. High throughput data comes to us from DNA and RNA sequencing, flow cytometry, metabolomics, molecular screens, and more. Although some accept the approach of compartmentalizing data generation and data analysis, we have found scientists feel empowered when they can both ask and answer their own biological questions [@Carey2018]. Yet, the standard undergraduate and graduate training in the biological sciences is typically insufficient for developing these advanced data analysis skills. In our experience performing microbiome research, it is more common to find exceptional bench scientists who are inexperienced at analyzing large data sets than to find the reverse. Of course this raises a challenge: How do we train bench scientists to effectively answer biological questions with these larger datasets?

The ever-growing ability to generate data and feeling of unease in analyzing the data are analogous to the struggles scientists also face with engaging the voluminous scientific literature. A common strategy for keeping up with the literature is participating in journal clubs, which involve group discussion of a pre-selected paper. These range from informal discussions to PowerPoint presentations and course credit. In addition to staying current on the literature, journal clubs help strengthen skills in critical thinking, communication, and integrating the literature [@Lonsdale2016]. Our research group has leveraged the similarity between the overwhelming nature of both scientific literature and data analysis to address the challenge of teaching reproducible data analysis practices to bench scientists. Over the past four years we have experimented with a Code Club model to improve data analysis skills in a community environment.

Our Code Club sessions are an hour long and alternate with Journal Club as the second part of weekly two hour lab meetings. Initially, the Code Club was used to review code from trainee projects. Instead of a presentation, the presenter would project their code onto a screen and the participants would go through the code, stating the logic behind each line. This approach emphasized the importance of code readability and gave beginners the opportunity to see the real-life, messy code of more experienced peers. Unfortunately, the format only allowed us to review a fraction of a project's code, making it difficult to integrate the programmer's logic across their full project. A major issue with this model was that sometimes beginners could not contribute to improving the code, and even when they could, more experienced group members would eventually dominate the discussion. This led to a lack of participation by beginners, who would sometimes mentally check out and resulted in an adversarial environment between the presenter and more experienced members. As a result, presenters were reluctant to offer to present again.

From these experiences, we began a collective conversation to improve our Code Club model. We have identified two successful approaches. The first is a more constructive version of a group code review. The presenter clearly states the problem they want to solve, breaks the participants into smaller groups, and then asks each group to solve the problem, or a portion of it. For example, someone may have an R script with a repeating chunk of code. The challenge for the session would be to convert the code chunk into a function to be called throughout the script to make it "DRY" (i.e. Don't Repeat Yourself [@Wilson2014]). The presenter leaves the session with several partial or working solutions to their problem and the importance of writing DRY code is reinforced. The second approach is a tutorial. The presenter introduces a new package or technique and assigns an activity to practice the new approach. For example, at one Code Club participants were given raw data and a finished plot. Paired participants were tasked with generating the plot from the data using either R syntax from the base language or the ggplot2 package. For this exercise, base R users had to use ggplot2 and vice versa. In either approach, the Code Club ends with a report back to the larger group describing the approach each pair took. Our Code Clubs typically have 7 to 10 participants, but the inherent "think-pair-share" strategy should allow it to be scaled to larger groups [@Kothiyal2014].

We continue to experiment with approaches for running Code Club. During the period of the Covid-19 pandemic in 2020 when many research labs were shuttered, PDS started hosting an open virtual weekly Code Club. Participants could engage the material posted to https://www.riffomonas.org/code_club/ via a Zoom meeting or asynchronously using the recorded sessions available on YouTube. Regardless of our approach to leading a Code Club, we have learned that it is critical for the presenter to clearly articulate their goals and facilitate participant engagement. Although some Code Club sessions may be more experimental than others, on the whole they are a critical tool to train bench scientists in reproducible data analysis practices. We have provided some examples of successful topics in Table 1. We have summarized the results of our own experimentation as Code Club presenters and participants into Ten Rules. The first 3 rules apply to all participants, Rules 4 through 8 are targeted to presenters, and the last two for participants.

@@

The selection of content and structure for each Code Club works best when it is democratic and distributed. If someone thinks a technique is worth learning and wants to teach it, they have that power. If they want to experiment with a different format, they are free to try it out. Members of the Code Club need to feel like they have the power to shape the direction of the group. If members are following these Rules, they will naturally reflect on the skills and interests of the other members in the group. For example, there is always turnover in a research group, making it important to revisit basic concepts to teach to new people and provide a refresher to others. Keeping an online log of previous topics or topics of interest for future sessions is a useful organizational tool. The group should also feel free to experiment with the format and incorporate group feedback by ending with a debrief discussing the pros and cons of new formats.

@@

Similar to a Journal Club, the presenter should give participants a few days (ideally a week) to prepare for the Code Club. Considering the compressed schedule described in Rule 5, asking participants to download materials beforehand is helpful to ensure a quick start. The presenter should provide the participants with instructions on how to install dependencies, download data, and get the initial code. This might also uncover weak points in the presenter's plan and enable them to ensure that the materials work as intended before the Code Club. We have found that using GitHub repositories for each Code Club can help make information, scripts, and data easily available to participants. Perhaps a first Code Club could introduce using git and GitHub to engage in collaborative coding. A lower barrier entry point for the presenter is posting their code, data, and information in a lab meeting-dedicated Slack channel or via email. Whatever method is used, the presenter should be sure to communicate the topic and necessary materials with the participants ahead of time.

@@

Participants want to learn topics that will either be useful to them or help their colleague (i.e. the presenter). Presenters should do their best to satisfy those motivations, whether it is through the relevancy of the concept or the data. It does not make sense to present a Code Club on downloading stock market data if it is not useful or interesting to the group. Similarly, participants should not be tasked with improving the presenter's code if the presenter has no intention of incorporating the suggestions. A list of packages or tasks that group members are interested in could help a presenter struggling to find a topic choose one that will make a rewarding Code Club.

@@

As suggested by Rule 7, a significant challenge to presenting at Code Club is selecting topics and activities that appeal to a critical mass of the participants. This is particularly difficult if participants have a wide range of coding experience, which can follow a turnover in group membership. In these circumstances it can be useful to cover commonly used tools but that might result in disinterest of more experienced participants. We have identified several strategies to overcome the challenges presented by participants at varying skill levels. Instead of letting participants form their own pairs, the presenter can select pairs of participants with either similar or differing skill levels, depending on their goals. Partnerships between those with similar skill levels requires the presenter to design appropriate activities for each skill level. We have found that commenting code is a good skill for beginners since it forces them to dissect and understand a code chunk line-by-line. It also reinforces the value of commenting as they develop their skills and independence. An advantage of forming partnership between people with disparate skill levels is that it is more likely for groups to provide the presenter with a diverse range of methods that achieve the same result. This approach to pairing also helps to graft new members that have emerging programming skills. Regardless of how partners are selected, consider asking the pairs to identify a navigator and a driver [@Bockeler_2020]. The driver types at the computer while the navigator tells them what to type, thus ensuring participation of both partners. Midway through the activity, the presenter can have the partners switch roles. Intentionally forming pairs can also engineer group interactions by avoiding potentially disruptive partnerships or pairing reliable role models with new group members.

@@

It is not possible to fully participate in a Journal Club discussion about a paper that the participant has not read. In that context, coming to Code Club without having installed a dependency is similar to asking a simple question about the Journal Club paper. Both instances show a lack of preparation. Just as a presenter must follow Rule 6 to provide materials ahead of time, participants must review the code in advance, download the data sets, install the necessary packages, and perhaps read up on the topic. If the Code Club is based on a paper or chapters in a data science book (e.g., [@Wickham_2016; @Schloss_2018; @Noble2009]), the participants should read them before the session and consider how they might incorporate the concepts into their own work.

@@

The most important rules are the first and last. Members of the Code Club need to feel comfortable with other group members and sufficiently empowered to try something new or ask for help. Aside from expanding our programming skills, we have noticed two other benefits that help create a positive work culture. First, we have intentionally interviewed postdoc candidates on the days we hold Code Club. We make it clear that they are not being assessed on their coding skills, but instead use it as an opportunity to see how the candidate interacts with other members of the research group. At the same time, the candidate can learn about the culture of the research group through active participation. Second, members of other research groups have integrated themselves into our Code Club to minimize the isolation they feel in growing their skills within smaller research groups. This speaks to both the broader need for Code Club and the likelihood of success when expanded to include a larger group of individuals with broader research interests. Ultimately, Code Club has improved the overall data analysis skills, community, and research success of our lab by empowering researchers to seek help from their colleagues.

@@

| **Title**             | **Description**                                    |
| ----------------- | ---------------------------------------------- |
| base vs. ggplot2   | Given input data and a figure, recreate the figure using R's base graphics or ggplot2 syntax   |
| Snakemake         | Given a bash script that contains an analysis pipeline, convert it to a Snakemake workflow (can also be done with GNU Make) |
| DRYing code       | Given script with repeated code, create functions to remove repetition   |
| mothur and Vegan	| Given a pairwise community dissimilarity matrix, compare communities using the `adonis` function in the Vegan R package  |
| tidy data         | Given a wide-formatted data table, convert it to a long, tidy-formatted data table using tools from R's tidyverse  |
| GitFlow           | Participants file and claim an issue to add their name to a README file in a GitHub-hosted repository and file a pull request to complete the issue |
| R with Google docs   | Scrape a Google docs workbook and clean the data to identify previous Code Club presenters |
| Develop an R Package  | Convert a lab member's collection of scripts into an R package over a series of sessions  |

>>>>

For most biologists, the ability to generate data has outpaced the ability to analyze those data. High throughput data comes to us from DNA and RNA sequencing, flow cytometry, metabolomics, molecular screens, and more. Although some accept the approach of compartmentalizing data generation and data analysis, we have found scientists feel empowered when they can both ask and answer their own biological questions [@Carey2018]. In our experience performing microbiome research, it is more common to find exceptional bench scientists who are inexperienced at analyzing large data sets than to find the reverse. Of course this raises a challenge: How do we train bench scientists to effectively answer biological questions with these larger datasets?

The standard undergraduate and graduate training in the biological sciences is typically insufficient for developing these skills [@Barone2017]. To fill this void, there has been growth in online data science training resources through companies including Data Camp (https://www.datacamp.com) and Codecademy (https://www.codecademy.com), massive open online courses (MOOCs), and in person workshops hosted by organizations including The Carpentries (https://carpentries.org). Each of these formats are highly popular among their audience (e.g. [@Wilson2016]); however, recent analyses of workshops suggests that they have little impact on the long term retention of the material that is taught [@Feldon2017]. This result should not be surprising. These learning formats ask participants to engage in massed practice; an approach that is known to be ineffective [@Carpenter2012; @Rohrer2015b; @Bud2010; @Karpicke2011]. In contrast, participants need regular opportunities to engage in repeated deliberate practice that encourage them to practice their retrieval and application skills of the material [@Rohrer2015a; @Karpicke2008]. There is clearly a need to develop a framework that can build upon material introduced in coursework and workshops and that can help researchers to stay up to date on the latest best practices, algorithms, and tools.

To address this need, we noted the similarity between the feeling of unease in analyzing the data and the struggles scientists face with engaging the voluminous scientific literature. A common strategy for keeping up with the literature is by participating in a journal club. Regardless of the discipline, a Journal Club involves group discussion of a pre-selected paper. These range from informal discussions to PowerPoint presentations and course credit. In addition to building upon material from traditional coursework and staying current on the literature, journal clubs help strengthen skills in critical thinking, communication, and integrating the literature [@Lonsdale2016]. Since most journal clubs occur on a regular schedule, they are effective because they incorporate repeated deliberate practices. With this model in mind, over the past four years we have experimented with creating a Code Club model with the goal of improving reproducible data analysis skills in a community environment.

Our Code Club sessions are an hour long and alternate with our lab's Journal Club as the second part of weekly two hour lab meetings. Initially, the Code Club was used to review code from trainee projects. Instead of a presentation, the presenter would project their code onto a screen and the participants would go through the code, stating the logic behind each line. This approach emphasized the importance of code readability and gave beginners the opportunity to see the real-life, messy code of more experienced peers. Unfortunately, the format only allowed us to review a fraction of a project's code, making it difficult to integrate the programmer's logic across their full project. A major issue with this model was that sometimes beginners could not contribute to improving the code, and even when they could, more experienced group members would eventually dominate the discussion. This led to a lack of participation by beginners, who would sometimes mentally withdraw, resulting in an adversarial environment between the presenter and more experienced members. As a result, presenters were reluctant to offer to present again.

From these experiences, we began a collective conversation to improve our Code Club model. We have identified two successful approaches. The first is a more constructive version of a group code review. The presenter clearly states the problem they want to solve, breaks the participants into smaller groups, and then asks each group to solve the problem, or a portion of it. For example, someone may have an R script with a repeating chunk of code. The challenge for the session would be to convert the code chunk into a function to be called throughout the script to make it "DRY" (i.e. Don't Repeat Yourself [@Wilson2014]). The presenter leaves the session with several partial or working solutions to their problem and the importance of writing DRY code is reinforced. The second approach is a tutorial. The presenter introduces a new package or technique and assigns an activity to practice the new approach. For example, at one Code Club participants were given raw data and a finished plot. Paired participants were tasked with generating the plot from the data using either R syntax from the base language or the ggplot2 package. For this exercise, base R users had to use ggplot2 and vice versa. In either approach, the Code Club ends with a report back to the larger group describing the approach each pair took. We generally find that preparing a Code Club session takes as much effort as preparing to present at a journal club. Our Code Clubs typically have 7 to 10 participants, but the inherent "think-pair-share" approach should allow it to be scaled to larger groups [@Kothiyal2014].

We continue to experiment with approaches for running Code Club. For example, during the period of the Covid-19 pandemic in 2020 when many research labs were shuttered, PDS started posting Code Club materials (https://www.riffomonas.org/code_club/) that individuals or research groups can use to practice their coding skills. During this time, we continued our group's Code Club sessions by making use of the breakout room and screen sharing features in Zoom. Regardless of our approach to leading a Code Club, we have learned that it is critical for the presenter to clearly articulate their goals and facilitate participant engagement. Although some Code Club sessions may be more experimental than others, on the whole they are a critical tool to train bench scientists in reproducible data analysis practices. We have provided some examples of successful topics in Table 1. We have summarized the results of our own experimentation as Code Club presenters and participants into Ten Rules. The first 3 rules apply to all participants, Rules 4 through 8 are targeted to presenters, and the last two for non-presenters. As the rules describe, this model can easily be adapted to groups of non-biologists and to groups with stronger background in computer science.

@@

The selection of content and structure for each Code Club works best when it is democratic and distributed. If someone thinks a technique is worth learning and wants to teach it, they have that power. If they want to experiment with a different format, they are free to try it out. Members of the Code Club need to feel like they have the power to shape the direction of the group. If members are following these Rules, they will naturally reflect on the skills and interests of the other members in the group. For example, there is always turnover in a research group, making it important to revisit basic concepts to teach to new people and provide a refresher to others. One successful experiment found that we could keep basic content interesting to more experienced group members by creating related problems that varied in their difficulty. The group should also feel free to experiment with the format and incorporate group feedback by ending with a debrief discussing the pros and cons of new formats.

@@

Similar to a Journal Club, the presenter should give participants a few days (ideally a week) to prepare for the Code Club. Considering the compressed schedule described in Rule 5, asking participants to download materials beforehand is helpful to ensure a quick start. The presenter should provide the participants with instructions on how to install dependencies, download data, and get the initial code. This might also uncover weak points in the presenter's plan and enable them to ensure that the materials work as intended before the Code Club. We have found that using GitHub repositories for each Code Club can help make information, scripts, and data easily available to participants. Although this can be convenient, introducing GitHub on top of the session's activities can impose a significant cognitive load and frustration to new group members as well as to the presenter if they are not comfortable with GitHub. Perhaps a first Code Club could introduce using git and GitHub to engage in collaborative coding. A lower barrier entry point for the presenter is posting their code, data, and information in a lab meeting-dedicated Slack channel or via email. Whatever method is used, the presenter should be sure to communicate the topic and necessary materials with the participants ahead of time.

@@

Participants want to learn topics that will either be useful to them or help their colleague (i.e. the presenter). Presenters should do their best to satisfy those motivations, whether it is through the relevancy of the concept or the data. It does not make sense to present a Code Club on downloading stock market data if it is not useful or interesting to the group. Similarly, participants should not be tasked with improving the presenter's code if the presenter has no intention of incorporating the suggestions. A list of packages or tasks that group members are interested in along with a log of previous topics could help a presenter struggling to find a topic choose one that will make a rewarding Code Club.

@@

As suggested by Rule 7, a significant challenge to presenting at Code Club is selecting topics and activities that appeal to a critical mass of the participants. This is particularly difficult if participants have a wide range of coding experience, which can follow a turnover in group membership. Beginners will benefit from sessions that cover fundamental concepts and functions. The benefits experienced by more experienced participants include the improved understanding of concepts that comes through teaching the concepts and breaking down problems into simpler elements. Furthermore, instead of focusing on core functions, the group could balance Code Club sessions that cover basics with those that cover methods and packages that are new to everyone in the group. We have also identified several strategies to overcome the challenges presented by participants at varying skill levels. Central to the Code Club format is the use of paired-programming [@Hanks2011]. Instead of letting participants form their own pairs, the presenter can select pairs of participants with either similar or differing skill levels, depending on the presenter's goals. Partnerships between those with similar skill levels requires the presenter to design appropriate activities for each skill level. We have found that commenting code is a good skill for beginners since it forces them to dissect and understand a code chunk line-by-line. It also reinforces the value of commenting as they develop their skills and independence. An advantage of forming partnership between people with disparate skill levels is that it is more likely for groups to provide the presenter with a diverse range of methods that achieve the same result. This approach to pairing also helps to graft new members that have emerging programming skills. Regardless of how partners are selected, consider asking the pairs to identify a navigator and a driver [@Bockeler_2020]. The driver types at the computer while the navigator tells them what to type, thus ensuring participation of both partners. Midway through the activity, the presenter can have the partners switch roles. Intentionally forming pairs can also engineer group interactions by avoiding potentially disruptive partnerships or pairing reliable role models with new group members.

@@

It is not possible to fully participate in a Journal Club discussion about a paper that the participant has not read. In that context, coming to Code Club without having installed a dependency is similar to asking a simple question about the Journal Club paper without first reading it. Both instances show a lack of preparation. Just as a presenter must follow Rule 6 to provide materials ahead of time, participants must review the code in advance, download the data sets, install the necessary packages, and perhaps read up on the topic. If the Code Club is based on a paper or chapters in a data science book (e.g., [@Wickham_2016; @Schloss_2018; @Noble2009]), the participants should read them before the session and consider how they might incorporate the concepts into their own work.

@@

The most important rules are the first and last. Members of the Code Club need to feel comfortable with other group members and sufficiently empowered to try something new or ask for help. Aside from expanding our programming skills, we have noticed two other benefits that help create a positive culture. First, we have intentionally interviewed postdoc candidates on the days we hold Code Club. We make it clear that they are not being assessed on their coding skills, but instead use it as an opportunity to see how the candidate interacts with other members of the research group. At the same time, the candidate can learn about the culture of the research group through active participation. Second, members of other research groups have integrated themselves into our Code Club to minimize the isolation they feel in growing their skills within smaller research groups. This speaks to both the broader need for Code Club and the likelihood of success when expanded to include a larger group of individuals with broader research interests. There is no reason that the Code Club format would not work with a group of people independent of a research group as long as the follow the Rules. Finding datasets and applications that are interesting to a critical mass of people is likely critical to starting and sustaining such a group. Ultimately, Code Club has improved the overall data analysis skills, community, and research success of our lab by empowering researchers to seek help from their colleagues.

@@

| **Title**             | **Description**                                    |
| ----------------- | ---------------------------------------------- |
| base vs. ggplot2   | Given input data and a figure, recreate the figure using R's base graphics or ggplot2 syntax   |
| Snakemake         | Given a bash script that contains an analysis pipeline, convert it to a Snakemake workflow (https://github.com/SchlossLab/snakemake_riffomonas_tutorial) |
| DRYing code       | Given script with repeated code, create functions to remove repetition   |
| Vegan R package	| Compare microbial communities using the `adonis` function in the Vegan R package (https://github.com/SchlossLab/Code_Review_42717)  |
| tidy data         | Given a wide-formatted data table, convert it to a long, tidy-formatted data table using tools from R's tidyverse  |
| GitFlow           | Participants file and claim an issue to add their name to a README file in a GitHub-hosted repository and file a pull request to complete the issue |
| googledocs4 R package | Scrape a Google docs workbook and clean the data to identify previous Code Club presenters |
| Develop an R Package  | Convert a lab member's collection of scripts into an R package over a series of sessions  |
| Documenting R code  | Use roxygen2 to supplement comments in R code to improve documentation (https://github.com/SchlossLab/documenting-R) |
| gganimate R package  |  Convert static plots generated with ggplot2 package into GIFs (https://github.com/SchlossLab/2020-04-12-CodeClub_PlotAnimation/) |
